Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-11-16 demystify.tex
updates

\section{Is an MS required for admission to PhD in CS?}\label{sec:msrequirement}
\sectioninfo{You do not need an MS to do PhD in CS.}

No. In fact, students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.  However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).

If you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But in general don't count finishing earlier just because you have an MS.

@@

\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:msrequirement} and \autoref{sec:time}).  In contrast, many other countries require having an MS degree before joining a PhD program.  This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).

>>>>

\section{Is an MS required for admission to PhD in CS in the US?}\label{sec:msrequirement}
\sectioninfo{You do not need an MS to do PhD in CS.}

No, not in the US (other countries often do though~\autoref{sec:non-us-differences}).  In the US, it is common to apply for a PhD program directly after your undergrad.  Most CS PhD programs are designed so that students can get MS degree ``along the way'' to PhD, e.g., after finishing the 2-year course work.

However, MS can help admission if it gives research experience or is from a more well-known school than your undergrad institution (\autoref{sec:your-school}).
Moreover, if you have an MS then some course work \emph{might be} transferred for course credits, which \emph{might} save some time. But in general don't count finishing earlier just because you have an MS.

@@

\emph{MS requirement and PhD duration}:  CS PhD programs in the US do not require an MS degree (as mentioned in \autoref{sec:time} and \autoref{sec:msrequirement}).  In contrast, many other countries require having an MS degree before joining a PhD program.  This means that US PhD programs are longer (5--7 years, 2 of which are coursework) than other countries (3--4 years, no coursework).
